On Devon Allen's Disqualification at the 2022 Track and Field World Championship

### 2023-04-18 Manuscript/manuscript.tex
minor changes

\jy{use the following structure and flow:
  1) A paragraph to tell the story;
  2) A paragraph on informal statistical analyses on the internet/social media;
  3) A paragraph on scholarly articles on reaction time in the literature;
  4) A paragraph on what's new: methods; results.
  5) A paragraph of roadmap.}

@@

A 2012 follow-up study to Komi's by \citet*{pilianidis2012start} found
significant differences between reaction times at several world championships
between 1997 and 2009 for the men's 110 meter hurdles.  Another important
conclusion was that for reaction times did not decrease over the twelve years
covered by the research in the study \citep{pilianidis2012start}.  This contradicts
what was found in a 2021 study that concluded the year of competition has a
significant effect on reaction times \citep*{zhang2021correlation}.  Their
analysis looked at data from the 2011-2019 World Championships across multiple
sprint events and concluded that from 2011 to 2019, reaction times got
faster. However, not  all studies have been in favor of a decrease in the
reaction time barrier. Brosnan in 2016 argues for increasing the reaction time 
barrier to 0.115 seconds, and cited evidence from European Championships and 
World Championships in his article \citep{brosnan2017effects}.  A study looking 
at the 2008 Olympics likewise supported the conclusion that athletes could not
react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
concluded that visual reaction times to the starting gun may be faster by 0.007
seconds than the IAAF method, which is from force on block starts 
\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
only by 0.01 seconds. It is also worth noting that one study looking at data 
from various IAAF meets found that the fastest reaction times for men were found
between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
because Allen was 27 years old at the 2022 Track and Field World Championship 
and in his physical peak.

@@

%4
\jy{I think you have one objective but two methods to approach it.}
One of the first objectives of the paper is to try to model the data, which is
the reaction times of every non-disqualification since the 1999 World Championships. 
The model we use is a linear model, following a gamma distribution with a random
effect, which will be referred to as the venue effect.  The venue effect is
reflective of how significant the changes in the reaction time is for each venue,
i.e. the venue effect is  responsible for the variation between years.  
Another objective of this paper is to compare data for athletes who have competed
at additional competitions to see how their reaction times compare from previous
competitions to this one. First we can look at the reaction times of athletes who
competed at the World Track and Field Championship in both 2019 and 2022 and
compare the times, and secondly we can look at how athletes performed compared
to how they did at their national/qualifiers meet. In order to reduce any chances
of athletes improving reaction time, we can also look at data from qualifying meets.
Almost every athlete had to qualify by proving they are elite at the national level, 
usually by coming in at least fourth. Thus, data was compiled on athletes who competed
at the 2022 World Track and Field Championships on what these athletes also ran 
at their country's national championship. If there was any argument that athletes
have improved their reaction time from 2019 to 2022, it should now be considered
nearly null, as the national competitions took place between June and July 2022 
and the World Championships were in August. The equipment used to measure reaction
time cannot be guaranteed to be the same from national to international
competition, but the time span between meets has been reduced considerably.

@@

At a size of $n=10,000,000$, the calculated probability of a reaction
time being below 0.1 seconds was as follows for the various different data sets:
\jy{label the table and reference to it; clean it up}
\begin{center}
  \begin{tabular}{c c} 
   \toprule
   Data Set & Probability RxnTime < 0.1 \\ 
   \midrule
   all finals gamma log & 0.000856 \\ 
   old finals gamma log & 0.000439 \\
   all pooled gamma log & 0.006865 \\
   old pooled gamma log & 0.005761 \\
   \bottomrule
   \label{fig:Sim_probability}
  \end{tabular}
  \end{center}

>>>>

\jy{use the following structure and flow:
  1) A paragraph to tell the story;
  2) A paragraph on informal statistical analysis on the internet/social media;
  3) A paragraph on scholarly articles on reaction time in the literature;
  4) A paragraph on what's new: methods; results.
  5) A paragraph of roadmap.}

@@

A 2012 follow-up study to Komi's by \citet*{pilianidis2012start} found
significant differences between reaction times at several world championships
between 1997 and 2009 for the men's 110 meter hurdles.  Another important
conclusion was that for reaction times did not decrease over the twelve years
covered by the research in the study \citep{pilianidis2012start}.  This contradicts
what was found in a 2021 study that concluded the year of competition has a
significant effect on reaction times \citep{zhang2021correlation}.  Their
analysis looked at data from the 2011-2019 World Championships across multiple
sprint events and concluded that from 2011 to 2019, reaction times got
faster. However, not  all studies have been in favor of a decrease in the
reaction time barrier. \citet*{brosnan2017effects} in 2016 argue for increasing 
the reaction time barrier to 0.115 seconds, and cited evidence from European 
Championships and World Championships in their article.  A study looking 
at the 2008 Olympics likewise supported the conclusion that athletes could not
react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
concluded that visual reaction times to the starting gun may be faster by 0.007
seconds than the IAAF method, which is from force on block starts 
\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
only by 0.01 seconds. It is also worth noting that one study looking at data 
from various IAAF meets found that the fastest reaction times for men were found
between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
because Allen was 27 years old at the 2022 Track and Field World Championship 
and in his physical peak.

@@

%4
\jy{I think you have one objective but two methods to approach it.}
The objective to this paper is to analyze the reaction times at the 2022
World Track and Field Championship to check for any abnormalities, and we approach
this two different ways.  Our first method is to try to model the data, which is
the reaction times of every non-disqualification since the 1999 World Championships. 
The model we use is a linear model, following a gamma distribution with a random
effect, which will be referred to as the venue effect.  The venue effect is
reflective of how significant the changes in the reaction time is for each venue,
i.e. the venue effect is  responsible for the variation between years.  
The second method of this paper is to compare data for athletes who have competed
at additional competitions to see how their reaction times compare from previous
competitions to this one. First we can look at the reaction times of athletes who
competed at the World Track and Field Championship in both 2019 and 2022 and
compare the times, and secondly we can look at how athletes performed compared
to how they did at their national/qualifiers meet. In order to reduce any chances
of athletes improving reaction time, we can also look at data from qualifying meets.
Almost every athlete had to qualify by proving they are elite at the national level, 
usually by coming in at least fourth. Thus, data was compiled on athletes who competed
at the 2022 World Track and Field Championships on what these athletes also ran 
at their country's national championship. If there was any argument that athletes
have improved their reaction time from 2019 to 2022, it should now be considered
nearly null, as the national competitions took place between June and July 2022 
and the World Championships were in August. The equipment used to measure reaction
time cannot be guaranteed to be the same from national to international
competition, but the time span between meets has been reduced considerably.

@@

At a size of $n=10,000,000$, the calculated probability of a reaction
time being below 0.1 seconds was as follows for the various different data sets:
\jy{label the table and reference to it; clean it up}
\begin{center}
  \begin{tabular}{c c} 
   \toprule
   Data Set & Probability RxnTime $<$ 0.1 \\ 
   \midrule
   all finals gamma log & 0.000856 \\ 
   old finals gamma log & 0.000439 \\
   all pooled gamma log & 0.006865 \\
   old pooled gamma log & 0.005761 \\
   \bottomrule
   \label{fig:Sim_probability}
  \end{tabular}
  \end{center}
